# Classifying Quality of Barbell Exercises from Sensor Data

## Executive Summary
In this analysis, I take sensor data collected during barbell exercises and build a Random Forest model to classify the quality of the exercise. After cleaning the data to strip non-predictor variables, I partition it into a train set (80% of data) and a validation set (20% of data). I build a Random Forest model on the training set and estimate the out of sample error on the validation set, finding an estimate of 0.57%. Finally, I predict the exercise quality on a separate test set of 20 samples.

## Analysis

### Import data
```{r}
data <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

### Load Libraries
```{r echo=FALSE}
library(caret)
library(randomForest)
```

### Clean Data
```{r}
originalData <- data
# Remove rows with newwindow == 'yes'
data <- data[data$new_window != "yes",]

# Remove columns with NA's
data <- data[, colSums(is.na(data)) == 0]

# Remove empty columns
data <- data[,colSums(data == "") == 0]

# Remove Timestamps and other non-predictors
data <- subset(data, select=-c(X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window, num_window))
```

### Split Data into Training and Validation set
```{r}
set.seed(2526)
inTrain <- createDataPartition(y=data$classe, p=0.8, list=FALSE)
training <- data[inTrain,]
validation <- data[-inTrain,]
```

### Train Random Forest
```{r cache=TRUE}
set.seed(443)
modFit.rf <- randomForest(classe ~ ., data=training, importance=TRUE)
```

### Predict on Validation dataset
```{r}
valPred.rf <- predict(modFit.rf, validation)
```

### Check accuracy of predictions on validation set
```{r}
table(valPred.rf, validation$classe)
correct <- valPred.rf == validation$classe
```

The accuracy of the model on the validation dataset is `r sum(correct) / length(correct)`, giving an estimated out of sample error rate of `r 1 - sum(correct)/length(correct)`


### Predict on Testing dataset
```{r echo=TRUE}
testPred.rf <- predict(modFit.rf, testing)
```


## Appendix


### Write predictions to files
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(testPred.rf)
```

